|

### 2020-11-22 nontex/install/windows/fn_installations_R.Rmd
additional installation guides

---
title: |
  | Data-Science Set-up
  | Python (Anaconda), R (Rstudio), Matlab and Latex (Texlive)
  | VScode, Atom, Sublime Pycharm
output:
  html_document:
    number_sections: true
    toc: yes
  html_notebook:
    number_sections: true
    toc: yes
  word_document:
    number_sections: true
  pdf_document:
    df_print: paged
    number_sections: true
    toc: yes
urlcolor: blue
always_allow_html: yes
---

Go back to [fan](http://fanwangecon.github.io/)'s [Tex4Econ and Miscellaneous Repository](http://fanwangecon.github.io/Tex4Econ/). This file contains various excessive details on installations.

@@

# Objective

Install various compilers and IDE:

1. Python (Anaconda, Pycharm + VSCode + Atom, Poetry)
2. R (VSCode + Atom)
3. Matlab
4. Latex and PDF (texlive, Okular, VSCode + Atom + Texstudio)

Test at the end:

1. Do sample *py* files work?
    + work from command line
    + work from atom (hydrogen), Pycharm, VSCode
2. Do sample *rmd* files work?
    + work from r-studio
    + work from atom (hydrogen), VSCode
3. Do tex templates compile?
    + tikz files, template files with bib
    + work from texstudio, VSCode live compile, Atom

Installation Plan outline:

1. Conda install (python)
2. Install Atom along with Hydrogen and other packages.

# Python Testing and Installation with Poetry and Docs

When to use Python:

1. Main use Python for Dynamic programing type tasks
2. All immediate data tasks should be done in R, including things like generating data structures. Also graph generation in R.
3. Certain basic tabulation summary statistics etc things done in Stata

Two types of python usage cases:

1. Small chunks of python code in RMD for tutorial/class files, combine files in book format as bookdown.
2. Python installable packages and modules created with poetry, push to github as pypi: Locally test specially functions in [poetry](https://python-poetry.org/docs/#installation) environment. Push to github and Pypi, and then install via pip for system wide use.

Installations:

1. Python package development in poetry virtual environment, install with needed dependencies.
	- This means, during python package development, do not install the packages needed directly in conda, but add as dependency in Poetry
2. General environment with pip installed packages and conda installed path discoverable pathon packages for rmd chunks.

## Poetry

Install Python via Anaconda first following other instructions. Possibly cleanly uninstall and reinstall. Use Poetry to build python packages and modules.

For development machines, install [Poetry](https://github.com/python-poetry/poetry) and nstall [sphinx](https://www.sphinx-doc.org/en/master/).

```{bash}
# poetry
curl -sSL https://raw.githubusercontent.com/python-poetry/poetry/master/get-poetry.py | python
poetry --version
# sphinx
pip install sphinx
```

Theme for sphinx
```{bash}
pip install sphinx-rtd-theme
pip install sphinx_bootstrap_theme
pip install nbsphinx --user
```

Initialize a Poetry-style Python Package and add Sphinx style docs. This creates the folder *~/pyfan* and initialize it with various:

```{bash}
# creates ~/pyfan as poetry packaging
cd ~
poetry new pyfan

# sphinix quickstart
cd ~/pyfan
mkdir docs
sphinx-quickstart
make html

# build empty package
rm -rf dist/ build/
poetry install
poetry build
poetry publish

# What is the virtual Env id
poetry env info
```

### Poetry Update Python Version

[poetry change python version to 3.x](https://stackoverflow.com/questions/60580113/poetry-change-python-version-to-3-x)

The existing poetry env needs to be deleted and re-created if need to update python version.

```{bash}
poetry env list 
```

1. Update python requirement in .toml to *python = "^3.8"*
2. delete current env: *poetry env remove python*
3. run: *poetry lock*
4. run: *poetry update*, *poetry install*, etc...

### Installating Poetry Created Packages

Normally install/test/use inside conda virtual env or poetry virtual env. Poetry would install with *poetry add* and *poetry install*. After seeing some package of interest on github etc, see if the package is on pypi. pip install from pypi if possible.

Local Install

```{bash}
pip install --no-index --find-links="C:/Users/fan/Py4Econ/dist/pyfan-0.0.0.900.tar.gz" pyfan --ignore-installed
```

Pypi Install

Open up conda prompt, pip install from base. Will install dependencies of package that are in the poetry env but not otherwise in python installation main path.

```{bash}
pip install git+https://github.com/fanwangecon/pyfan.git#egg=pyfan
pip install pyfan --ignore-installed
pip show pyfan
```

Sometimes, someone ports a package and adds certain things that I need, so need to install from git a package whose modified version is not on pypi, then use git+https

#### What are egg, tar.gz, whl?

poetry creates tar.gz and whl in the dist folder, what are they?

- tarballs: tar.gz

### Developing and testing new tools Pycharm + Poetry

When tesing and developing tools, do so in the poetry virtual environment for the package in which testing is taking place.

After a package has been created and poetry built and poetry publish:

1. Type in *poetry env info*, and copy for example path: *C:/Users/fan/AppData/Local/pypoetry/Cache/virtualenvs/py4econ-6xnVtByB-py3.7*, to pycharm, so that can test module in pycharm using the poetry generated isolated packaging testing env.
2. In Pycharm: Settings (ctrl+alt+s), Project, Python interpreter.
3. Click on the cog button, "Add" > Right menu, "Virtual Environement" > Existing environement > Three dots > paste the path and point it to the python.exe in Scripts

```{bash}
poetry env info
```

When a particular tool is needed to build this new module inside package, add the tool via poetry add as a dependency. If a module requires another dependency: *poetry add numpy*, if adding numpy. This will install in the virtual env *numpy*. When packages are added, they are automatically installed. When the the *pyfan* package repo is downloaded somewhere, the dependencies might not be installed yet, and *poetry install* would install them.

```{bash}
poetry add numpy
```

#### Development for pyfan

pyfan is the general python tools package for various projects, not economics related directly.

Poetry added dependencies are in */pyfan/pyproject.toml*, Poetry add:

```{bash}
cd "C:/Users/fan/pyfan"
cd ~/pyfan
poetry env info
```

All installations:

```{bash}
# pyfan
poetry remove python-frontmatter pyyaml
poetry add python-frontmatter pyyaml
poetry add numpy
poetry add scipy
poetry add cython

# Compuational and Stats
poetry add statsmodels
poetry add numba
poetry add sklearn

# visualize
poetry add matplotlib
poetry add seaborn

# pyecon
poetry add pyfan

# aws
poetry add boto3
```

Poetry modules added over time:

```{bash}
# 2020-04
poetry add pyyaml
poetry add python-frontmatter
```

After modules are added and new programs developed

[sphinx readthedocx](https://samnicholls.net/2016/06/15/how-to-sphinx-readthedocs/)

SEE BELOW *pyfan/pyecon workflow and files added*:

```{bash}
cd ~/pyfan
cd "C:/Users/fan/pyfan"
cd "C:/Users/fan/pyecon"
poetry install

<!-- rm -rf dist/ build/ docs/
mkdir docs -->
cd doc/
rm -rf _build/ auto_examples/ gen_modules/backreferences gen_modules/generators
sphinx-build -b html ../ _build
<!-- make html -->
sphinx-build -b latexpdf ../ _build
make latexpdf

poetry build
poetry publish
```

SEE BELOW *pyfan/pyecon workflow and files added*, on some local computer install pyfan:

```{bash}
pip show pyfan
# conda activate pyfan
pip install pyfan
pip install git+https://github.com/fanwangecon/pyfan.git#egg=pyfan
```

```{bash}
ipython
import pyfan.util.rmd.bookdownparse as bookdownparse
?bookdownparse.fs_yml2readme
```

##### Uninstall

uninstall packages in the overall environment
```{bash}
# list all installed packages
pip list
pip show pyyaml
pip show python-frontmatter
pip show pyfan
# unstall packages
pip uninstall pyyaml
pip uninstall python-frontmatter
```

##### pyfan/pyecon workflow and files added

New python functions are added, or new Rmd files are added with python exampls.

List of to dos and checks.

New rmd files:

1. Edit yaml to include functions used
  + Go over each part, add to short desc
  + In each section, add code usage
2. Link to py functions if implemented there
3. Run file internally
4. Add new files to *_bookdown.yml*
5. Run pyfan rmd compile
6. Overall bookdown check pdf and html

New py functions:

1. Document with function name, parameters and returns
2. update version number
3. Test run the function file's main 
4. Local sphnix text etc
5. Poetry publish
6. Git push 
7. Check Docs on readthedocs
8. Uninstall and install locally
9. Uninstall and install in wk_pyfan envir

```{bash}
conda activate base

<!-- to update all dependencies -->
poetry update

cd "C:/Users/fan/pyfan/"
cd "C:/Users/fan/pyecon/"
poetry install

cd "C:/Users/fan/pyfan/doc/"
cd "C:/Users/fan/pyecon/doc"
rm -rf _build/ auto_examples/ gen_modules/backreferences gen_modules/generators
make html

<!-- Check Results in: C:\Users\fan\pyfan\doc\_build\html\index.html -->

cd "C:/Users/fan/pyfan/"
cd "C:/Users/fan/pyecon/"
poetry build
poetry publish
```

pip install works inside conda env. Uninstall first. Then install pyfan in conda env using pip. With user option, seems to install system-wide.

Note that if some of the conda env use different python versions than base, then when base programs are updated, will not update in env.

Might need to update this in all sub-directories

```{bash}
<!-- local install, this does not install in conda env -->
cd "C:/Users/fan/pyfan/"
cd "C:/Users/fan/pyecon/"
python setup.py install --user
<!-- after installing show appear under pip: pip show pyfan, with latest version ID -->
<!-- after installing this does not appear on conda list: conda list -n base, but in python, import pyfan as pyfan works -->
<!-- help('modules') in python to see if installed-->

pip show pyfan
<!-- pip install pyfan -->
pip uninstall pyfan -y
pip install pyfan
pip install git+https://github.com/fanwangecon/pyfan.git#egg=pyfan

conda activate wk_pyfan
conda activate wk_cgefi

<!-- if pyfan already installed, but not in conda env, uninstall and installing puts it in conda env -->
pip uninstall pyfan -y
pip install pyfan
<!-- --user installs it for outside of env -->
pip install pyfan --user

pip install git+https://github.com/fanwangecon/pyfan.git#egg=pyfan
conda deactivate
```

## sphinx and readthedocs

readthedocs is logged in via github account. Automatically recognizes github repositories, and can import repositories to generate documentation that it can.

1. install sphinx

```{bash}
pip install sphinx
cd ~/Py4Econ
mkdir docs
sphinx-quickstart
make html
```

2. install sphinx packages

- [phinx-gallery](https://sphinx-gallery.github.io/stable/index.html)

```{bash}
pip install sphinx-gallery
```

### autodoc

[Documenting Python Packages with Sphinx and ReadTheDocs](https://brendanhasz.github.io/2019/01/05/sphinx.html)

The Napoleon extension for Sphinx allows for NumPy/Google style docstrings instead of using the hard-to-read reStructuredText in your docstrings

> The autodoc extension for sphinx can automatically generate API reference doc pages from the docstrings in your python code. Python docstrings are string literals which occur immediately after function or class definitions. They’re treated as comments, and it’s customary to document a function or class in its docstring. The autodoc extension allows you to include reStructuredText in your docstrings, and will build an API reference automatically for your module from the docstrings, while allowing for further customization.

To enable the autodoc extension, in your conf.py file, add the string 'sphinx.ext.autodoc' to the extensions list (this should have already been done if you answered yes to enabling the autodoc extension during sphinx-quickstart).

```{python}
# If extensions (or modules to document with autodoc) are in another directory,
# add these directories to sys.path here. If the directory is relative to the
# documentation root, use os.path.abspath to make it absolute, like shown here.
#
# import os
# import sys
# sys.path.insert(0, os.path.abspath('.'))
```

After enabling autodoc, in the documentation’s .rst files, you can use the automodule directive:

.. automodule:: package_name.module
   :members:

.. autosummary::
   :toctree: _autosummary

   fantools.utility
   fantools.utility.inout
   fantools.utility.inout.bookdownparse
   fantools.utility.inout.rmdparse
   fantools.utility.timer

### rst

= for page titles,
- for sections,
^ for subsections, and
" for subsubsections.

Bulleted lists use asterisks, numbered lists use numbers, and nesting is done with indentation, like Markdown (though nested lists need to be surrounded by blank lines):

* Bulleted
* list

  * nested
  * list

* con't

1. Numbered
2. list

``Code sample``

For hyperlinks, use:

`Link text <http://www.link-address.com>`_

Commented lines begin with two periods and a space (like a directive w/o the double colons):

.. I'm a comment!
And multiline comments are also like directives w/o the colons:

..
   I am
   a multiline

   comment

## Python, Poetry, Docs, Git

Start from without a package. Create two packages: pyecon, pyfan

1. generate poetry repos first, before adding in .git with pull_repos
2.

```{bash}
cd ~
poetry new pyfan

cd ~
poetry new pyecon

base ~/base/bin/pull_repos

cd ~/pyfan
cd "C:/Users/fan/pyfan"
mkdir docs
<!-- change release number conf.py and .toml -->
<!-- sphinx-quickstart -->
make html
poetry build
poetry publish

cd ~/pyecon
cd "C:/Users/fan/pyecon"
mkdir docs
poetry build
poetry publish
sphinx-quickstart
make html

poetry env info
<!-- get virtual env info  -->
```

install git locally so can use outside of env

pip install git+https://github.com/fanwangecon/pyfan.git#egg=pyfan
pip show pyfan

> Name: pyfan
> Version: 0.1.2
> Summary:
> Home-page: None
> Author: Fan Wang
> Author-email: wangfanbsg75@live.com
> License: None
> Location: c:\programdata\anaconda3\lib\site-packages
> Requires:
> Required-by:

```{python}
import pyfan.util.rmd.rmdparse as ft_rmdparse
ft_rmdparse.fs_rmd_yml_parse()

import pyfan.util.rmd.bookdownparse as ft_bookdownparse
sfc_prj = 'R4Econ'
sph_prj = 'C:/Users/fan/R4Econ/'
spn_prj_bookdown_yml = '_bookdown.yml'
spn_prj_readme_toc = 'README_toc_three.md'

ft_bookdownparse.fs_yml2readme(sfc_prj=sfc_prj, sph_prj=sph_prj,
                          spn_prj_bookdown_yml=spn_prj_bookdown_yml,
                          spn_prj_readme_toc=spn_prj_readme_toc)
```

# Python and Anaconda Installation

Use conda across platforms, so that locally on windows and ubuntu and remotely on aws, can have the same software setup environment.

- Search for Anaconda Prompt, right click, choose run as administrator.
- Check software versions.

```{bash}
conda list anaconda
python -V
```

## Anaconda Install

### First Time Conda Install

[Download Anaconda Python 3](https://www.anaconda.com/distribution/) and install for all users. Afterwards, Anaconda does not automatically get added to Windows Path. Need to use Anaconda Prompt to access programs. To access Anaconda packages from windows prompt, from git bash, from R, etc, need to Add Anaconda to Windows Path.

To install conda in Linux/Debian, follow [these instructions](https://fanwangecon.github.io/Tex4Econ/nontex/install/linux/fn_ubuntu.html). Overall, installations are very similar.

In command/anaconda prompt:

```{bash}
# To remove conda Fully
rm -rf ~/anaconda3

where anaconda
# C:/ProgramData/Anaconda3/Scripts/anaconda.exe
where python
# C:/ProgramData/Anaconda3/python.exe
where jupyter
# C:/ProgramData/Anaconda3/Scripts/jupyter.exe
where jupyter-kernelspec
# if R installed already
where r

###############################
# Add these to Windows PATH:
###############################
# C:\ProgramData\Anaconda3
# C:\ProgramData\Anaconda3\Scripts
# C:\ProgramData\Anaconda3\Scripts\bin
```

To Add Anaconda to Path, In Windows
1. Search for: Environment Variables
2. Edit Environment Variables
3. Add new to Path (lower half):
    - C:/ProgramData/Anaconda3/
    - C:/ProgramData/Anaconda3/Scripts/
    - C:/ProgramData/Anaconda3/Scripts/bin        
4. Now open up regular windows command Prompt, Type in:
    - conda --version
    - also Close and Open up Git Bash: conda --version

### Conda Update, Install, Channels

Conda update python version

```{bash}
conda install -c anaconda python=3.7
conda update --all
```

Open up Anaconda Navigator, it will update navigator automatically. If there are errors, might have to clean first.

There are different channels/repositories from which packages could be installed: *conda*, *conda-forge*, for example. Can change where to look for packages first. Many packages are in multiple channels. Conda-forge often have mroe recent packages.

```{bash}
# if there are bugs
# conda clean --packages
# use conda-forge as main channel, more updated packages
conda config --get channels
conda config --add channels conda-forge
conda config --get channels
# remove conda-forge from channels (do so for main env install)
conda config --remove channels conda-forge
# normal update
# to show all currently installed: conda list -n base
conda update --all
```

### Conda Environments

For testing installed packages. Note that for development, using poetry for testing package dependencies for package development, and associate pycharm for projects with poetry virtual environments. The idea is that there might be dependency problems etc for packages depending on where/when they were made, might not work in the overall conda environment where pyfan and pyecon, my main tools are installed. Some specific task might only need one particular package. And when working on these, use a conda environment. When this works, could incorporate whatever I am working on into a pyfan, pyecon package along with other packages.

```{bash}
conda list env
conda env remove -n test_package_name
conda create -n test_package_name -y
conda activate test_package_name
conda list
conda deactivate

conda install pip

# see if package exists on pip and install from pip
pip install smop
```
#### Conda Base

```{base}
conda install -c anaconda boto3 -y

```

#### Conda Perl Environment

To run some perl tools in an isolated environment:

1. [latexpand](https://tex.stackexchange.com/a/21840/109818)

```{bash}
# Set up
conda deactivate
conda list env
conda env remove -n wk_perl
conda create -n wk_perl -y
conda activate wk_perl
conda install -c anaconda perl -y
conda list
conda deactivate

# intall packages in virtual env
cd "C:\Users\fan\.conda\envs\wk_perl\"
cd "C:\ProgramData\Anaconda3\envs\wk_perl\"
git clone https://gitlab.com/latexpand/latexpand.git

# use latexpand
conda activate wk_perl
cd "C:\Users\fan\Documents\Dropbox (UH-ECON)\Project Emily Minority Survey\Drafts_RR\OverleafChinaSchoolClosure"
perl "C:\Users\fan\.conda\envs\wk_perl\latexpand\latexpand" draft_edcc_submit_rr.tex > draft_edcc_submit_rr_flat.tex
```

#### Conda Jinja Environment

For testing and packaging latex files. In particular, single folder latex files for final paper submissions.

```{r}
# Set up
conda deactivate
conda list env
conda env remove -n wk_jinja
conda create -n wk_jinja -y
conda activate wk_jinja
conda install -c anaconda jinja2
conda install -c anaconda pyyaml

# use yaml2tex
cd "C:\Users\fan\Documents\Dropbox (UH-ECON)\repos\fanwangecon.github.io\_cv\templating\"
python yaml2tex.py
```

#### Conda Jekyll Environment (Ruby)

For testing github pages sites. On windows, install ruby using [windows-installer](https://rubyinstaller.org/downloads/). Windows choose install option 3, install toolchain.

```{r}
# Set up
conda deactivate
conda list env
conda env remove -n wk_ruby
conda create -n wk_ruby -y
conda activate wk_ruby

ruby -v
gem -v 
gcc -v
g++ -v 
make -v

gem install jekyll bundler

# use yaml2tex
cd "C:\Users\fan\Documents\Dropbox (UH-ECON)\repos\fanwangecon.github.io\_cv\templating\"
python yaml2tex.py
```

#### Conda wk_pyfan bookdown

revealjs, decktape to pdf conversions

```{r}
conda update -n base -c defaults conda

# Set up
conda deactivate
conda list env
conda env remove -n wk_pyfan
conda create -n wk_pyfan -y python=3.8
conda activate wk_pyfan

# add channel
# conda config --env --add channels conda-forge
conda config --get channels
conda config --get channels --env

# yaml
# conda install python-frontmatter -y
pip install python-frontmatter
conda install pyyaml -y

# pyfan install 
pip install pyfan

# numpy etc install
conda install -c anaconda numpy
# https://github.com/numpy/numpy/issues/17726
# https://github.com/numpy/numpy/wiki/FMod-Bug-on-Windows
# pip install numpy==1.19.3
# conda install -c anaconda cython

# matplotlib install
# conda install -c anaconda matplotlib -y
# python -m pip install -U matplotlib

# to test aws
conda install boto3 -y

# environment data
conda install cdsapi -y
conda install -c conda-forge eccodes -y
conda install -c conda-forge cfgrib

```

#### Conda ECMWF Envir

```

conda deactivate
conda list env
conda env remove -n wk_ecmwf
conda create -n wk_ecmwf -y
conda activate wk_ecmwf

conda install -c conda-forge cdsapi -y
conda install -c conda-forge eccodes -y
conda install -c conda-forge cfgrib -y
conda install -c conda-forge xarray -y
conda install -c conda-forge netcdf4 -y

```

#### Conda ECMWF Envir Python 2

```

conda deactivate
conda list env
conda env remove -n wk_ecmwf_py2
conda create -n wk_ecmwf_py2 python=2.7 -y 
conda activate wk_ecmwf_py2

python --version
```

#### Conda CGE Environment

Installing pip installed/updated python version from 3.7 to 3.8.

```
conda deactivate
conda list env
conda env remove -n wk_cgefi
conda create -n wk_cgefi -y
conda activate wk_cgefi

# Install External Tools
conda install -c anaconda pip -y

# Cloud
conda install -c anaconda boto3 -y

# Python tools
conda install -c anaconda pytest -y

# Install Own Tools
pip install pyfan

# Scientific Computing
conda install -c anaconda numpy -y
conda install -c anaconda statsmodels -y
conda install -c conda-forge interpolation -y
conda install -c anaconda pandas -y
conda install -c anaconda scikit-learn
conda install -c anaconda cython

# Graphing
conda install -c anaconda matplotlib -y
conda install -c anaconda pillow -y

```

#### Conda AWS Environment

AWS environment and set-up on local computer

```
conda deactivate
conda list env
conda env remove -n wk_aws
conda create -n wk_aws -y
conda activate wk_aws

install docker
```

#### Conda Bookdown for pyfan

To run some perl tools in an isolated environment:

1. [latexpand](https://tex.stackexchange.com/a/21840/109818)

```{bash}
# Set up
conda deactivate
conda list env
conda env remove -n wk_perl
conda create -n wk_perl -y
conda activate wk_perl
conda install -c anaconda perl -y
conda list
conda deactivate

# intall packages in virtual env
cd "C:\Users\fan\.conda\envs\wk_perl\"
cd "C:\ProgramData\Anaconda3\envs\wk_perl\"
git clone https://gitlab.com/latexpand/latexpand.git

# use latexpand
conda activate wk_perl
cd "C:\Users\fan\Documents\Dropbox (UH-ECON)\Project Emily Minority Survey\Drafts_RR\OverleafChinaSchoolClosure"
perl "C:\Users\fan\.conda\envs\wk_perl\latexpand\latexpand" draft_edcc_submit_rr.tex > draft_edcc_submit_rr_flat.tex
```
## Essential Programs to Install

Install Packages, note that some packages are dependencies for fantools and fanecon packages, they would be installed jointly with those packages.

```{bash}
pip install --user --upgrade setuptools wheel
pip install sphinx
pip install fantools --ignore-installed
```

Previously:

```{bash}
# install additional packages
# linux utilites
# conda install -y m2-base
# stats and graphs
# conda install -y statsmodels datashape seaborn
# various
# conda install -c conda-forge -y interpolation awscli
# aws communicate
# conda install -c anaconda -y boto3
```

## Fist test python under command line:

```{bash}
# windows
python "C:/Users/fan/PyFan/ProjectSupport/Testing/Numpy/Functions.py"
python "C:/Users/fan/PyFan/ProjectSupport/graph/subplot.py"
# linux
python ~/PyFan/ProjectSupport/Testing/Numpy/Functions.py
python ~/PyFan/ProjectSupport/graph/subplot.py
```

## Numpydoc

```{bash}
Returns:
    None

```

# R Installation

R could be installed from Anaconda or directly on windows outside of the Anaconda directories. When installing from within Anaconda, could create R environment that are isolated from the rest of the computer. The R environment could be updated and deleted without disturbing dependencies in conda main or the rest of the computer.

The computer could have multiple R installations. Several in different conda R environment, and also several under windows/linux primary main user R directories. Type *which R/where R* to see if you open up R from command line at the moment, which directory's R will be used. Inside Conda R environment, it will be a different R if there is a different R version there.

@@

If R was installed in R enviornments, just delete the environment. Otherwise, use system's uninstaller. Before that, from terminal/command-prompt, type *R*, and *.libPaths()* to find paths. Do so inside conda main, outside of conda main, inside different environments, find all paths. After uninstaller finishes, check in the libPath folders to see if there are still stuff there, delete all, delete the folders.

```{r}
.libPath()
# C:/Users/fan/Documents/R/win-library/3.6
# C:/Program Files/R/R-3.6.1/library
```

For Linux and for unsintalling inside conda:

```{bash}
# Exit Conda
conda deactivate
# where is R installed outside of Conda
which R
# /usr/bin/R
# To remove all
sudo apt-get remove r-base
sudo apt-get remove r-base-core

# Inside Conda base
conda activate
# Conda r_env
conda activate r_envr
# Where is it installed?
which R
# /home/wangfanbsg75/anaconda3/bin/R
conda uninstall r-base
```

## Install R (outside of Conda)

Need to install R. Then need to install also an editor/IDE for R. Rstudio is the dominant R IDE. VSCode and Atom also works well, especially VSCode as editors. Dramatically better experience for writing scripts outside of Rstudio which feels unwieldy and slow especially compared to VSCode.

Overall plan:

1. [download R](https://cloud.r-project.org/)
    + for debian: [Johannes Ranke](https://cran.r-project.org/bin/linux/debian/). For Linux/Debian installation, crucial to update the *source.list* to include sources that have more recent versions of R. If not, will get very old R versions that is not compatible with many packages.
    + add R to path for Windows. In Windows Path, add for example: *C:/Program Files/R/R-3.6.2/bin/x64/* and *C:/Rtools/bin*
2. [download R-studio](https://rstudio.com/products/rstudio/download/)
3. Open R-studio and auto-detect R
4. Install additional packages (from commandline)
    + various R data-science packages have dependencies, and additional not-R programs might need to be installed. Install them as needed, for example:  *libcurl4-openssl-dev* and *libssl-dev* from command prompt first inside Debian.

### Linux R Install

For linux/Debian, to Install latest R:

```{bash}
# Go to get latesdebian latest r sources.list
cat /etc/apt/sources.list
# Install this First (should already be installed)
sudo apt install dirmngr

# Debian R is maintained by Johannes Ranke, copied from https://cran.r-project.org/bin/linux/debian/:
apt-key adv --keyserver keys.gnupg.net --recv-key 'E19F5F87128899B192B1A2C2AD5F960A256A04AF'
# Add to source.list, for debian stretch (9)
# sudo su added for security issue as super-user
sudo su -c "sudo echo 'deb http://cloud.r-project.org/bin/linux/debian stretch-cran35/' >> /etc/apt/sources.list"
# if added wrong lines, delete 3rd line
sudo sed '3d' /etc/apt/sources.list

# Update and Install R, should say updated from cloud.r
sudo apt-get update
sudo apt-get install r-base r-base-dev

# Also install these, otherwise r-packages do not install
# libxml2 seems need for tidymodels
sudo apt-get install libcurl4-openssl-dev
sudo apt-get install libssl-dev
sudo apt-get install libxml2-dev
```

### Update R

@@

### Install Packages for R

What are the packages I need in R:

**Use Multiple Languages**:

- [reticulate](https://github.com/rstudio/reticulate)
    + install directly from: devtools::install_github("rstudio/reticulate")

**Visualization, Packaging Etc**:

- [bookdown](https://bookdown.org/)
- [formatR](https://cran.r-project.org/web/packages/formatR/index.html): to tidy up rmd chunks

From command prompt, enter R by typeing *R*, and install packages:

# Install Packaging tools
install.packages(c("devtools", "pkgdown", "roxygen2", "bookdown", "knitr", "kableExtra", "formatR"))

# Install Statistics models
install.packages(c("AER", "minpack.lm"))

# Personal Packages
devtools::install_github("fanwangecon/REconTools")
devtools::install_github("fanwangecon/PrjOptiAlloc")
```
```{r}
# Can enter R from Command Prompt Conda ENV, much faster than from r-studio
# Install R-tools, RTOOLS is for WINDOWS
# install.packages("installr")
# library("installr")
# install.Rtools()
# c:/Rtools/bin;
# c:/Rtools/mingw_32/bin;
# c:/Rtools/mingw_64/bin;

# Install RTools First!
https://cran.r-project.org/bin/windows/Rtools/

# main packagevps
# tidymodel installation on linux difficult
# install.packages(c("tidyverse", "tidymodels", "tidyr"))
# development packages
# install.packages(c("devtools", "IRkernel", "pkgdown", "roxygen2", "kableExtra"))
# other packages
install.packages(c("AER", "minpack.lm"))
# install.packages(c("AER", "minpack.lm", "knitr", "matlab"))

# for working with other tools
# install.packages(c("reticulate"))

# Install my package, need to prepare package to work for ubuntu
# if install does not work, load locally
devtools::install_github("fanwangecon/REconTools")

# Present
install.packages("revealjs")
# Kernel
# install.packages(c('repr', 'IRdisplay', 'evaluate', 'crayon', 'pbdZMQ', 'devtools', 'uuid', 'digest'))
# IRkernel::installspec()
```

## Install R (inside Conda)

Key packages are generally available in conda's default channel (official distribution) and also the more frequently updated conda-forge channel. Prioritize conda-forge to get the latest packages. After adding *conda-forge* to channels, that will be prioritized, so when creating a new environment, conda will install first from conda-forge if package exists there. Try the *r_env* generation line with and without first adding *conda-forge* to channel, and see that the packages installed will have different versions, reflecting versions in the default channel and in the *conda-forge* channels.

### Set up Conda R environment

Create a Conda Environment called *r_env*, and update its channels.

- **Location**: Installed environments are shown appear in the *envs* folder of *C:/ProgramData/Anaconda3/envs/*. Can be easily deleted without disrupting the main base installation for conda. This is very important. During installation, could easily run into problems, and need to re-install. Much better to only re-install a folder. Note that when R is installed in *envs*, it will not show up under add or remove programs for uninstallation from there, has to be uninstalled, deleted directly here.
- **Use only in Env**: Note that if we installed r inside r_env, if we type *r* under the base environment, we can not enter r. We can only enter r within r_env. With PATH properly set-up, this happens under Conda Prompt, Windows Prompt, etc.
- **Multiple Pythons**: With the above installation for R, one benefit is that if R requires a different version of Python than what the base environment uses, the r_env could have a different python version. So type *python* in the base environment as well as inside the *r_env*. This also means potentially there could be duplicated installations I think. Look at the python versions under *conda list* in the base and in the r-environment.
- **File Size**: Note that this creates a large installation folder, *C:/ProgramData/Anaconda3/envs/r_env/*, without additional packages, is 1.3 GB in size

```{bash}
# outside of conda, install some depencies
sudo apt-get install libcurl4-openssl-dev
sudo apt-get install libssl-dev
sudo apt-get install libxml2-dev

# install key r packages within a r-environment
conda env remove -n r_env
# start empty environment
conda create -n r_env
# activate env and check, nothing yet
conda activate r_env
conda list

# check channel and add env specific channel
conda config --get channels
# to get latest r, use conda-forge, which has more recent Rs
# conda config --env --add channels conda-forge
# channels for base
conda config --get channels
# channels for cur env
conda config --get channels --env

# see all installed environments
conda env list
# activate an environment to use it, if in base, r does not exist, isolated in r_env
conda activate r_env
# see packages in environment
conda list
# to quit
conda deactivate
```

### Install R and Some Packages in Conda

Either inside a R specific environment, or outside of it, now do additional installations for R and also R packages.

During installation, the R programs might require downgrading other programs in the Conda environment that Python for example also uses. That is why potentially it is safer to install in R environment. However, when installing in R enviornment, certain packages could have issues if installer does not realize it is in an environment with special folder structure, so installation could fail.

```{bash}
# install files
conda install -n r_env -c r r r-essentials r-tidyverse r-tidymodels -y

# trouble with igraph installation, install here from conda
# conda install -n r_env r-essentials r-base r-tidyr r-tidyverse r-devtools r-irkernel r-pkgdown r-roxygen2
conda install -n r_env -c r r-igraph
conda install -n r_env -c conda-forge igraph
conda install -n r_env -c r r-rstantools r-rstan r-rstanarm -y
```

## Install Additional Packages inside the R

These are packages that I use, these should be installed inside the environment. Or they could be installed later from inside r-studio. Installing inside r-studio is a lot better. *tidymodel* has installation issues sometimes.

In *r_env*, type in *R*, and then install from inside terminal's R. For conda env, devtools will be installed and will install packages inside *envs/r_env/Lib/R*.

```{r}
# enter into env
# conda activate r_env
# enter R or type in rstudio
# sicne R is only install inside r_env, has to enter rstudio from r_env
# R
# rstudio

# main packagevps
# tidymodel installation on linux difficult
install.packages(c("rstantools", "rstan", "rstanarm"))
install.packages(c("tidyr", "tidyverse", "tidymodels"))

>>>>

---
title: |
  | R Installation and Update
output:
  html_document:
    number_sections: true
    toc: yes
  html_notebook:
    number_sections: true
    toc: yes
  word_document:
    number_sections: true
  pdf_document:
    df_print: paged
    number_sections: true
    toc: yes
urlcolor: blue
always_allow_html: yes
---

Go back to [fan](http://fanwangecon.github.io/)'s [Tex4Econ and Miscellaneous Repository](http://fanwangecon.github.io/Tex4Econ/).

@@

# R Installation and Set-Up

Install R and Update R

@@

Uninstall R, RStudio and RTools from Windows "Programs and Features" menu

If R was installed in a virtual environment, delete the environment. Otherwise, use system's uninstaller. Before that, from terminal/command-prompt, type *R*, and *.libPaths()* to find paths. Do so inside conda main, outside of conda main, inside different environments, find all paths. After uninstaller finishes, check in the libPath folders to see if there are still stuff there, delete all, delete the folders.

```{r}
ls_spn_paths <- .libPaths()
print(ls_spn_paths)
# C:/Users/fan/Documents/R/win-library/3.6
# C:/Program Files/R/R-3.6.1/library
```

## Install R First Time

1. [download R](https://cloud.r-project.org/)
    + for debian: [Johannes Ranke](https://cran.r-project.org/bin/linux/debian/). For Linux/Debian installation, crucial to update the *source.list* to include sources that have more recent versions of R. If not, will get very old R versions that is not compatible with many packages.
    + add R to path for Windows. In Windows Path, add for example: *C:/Program Files/R/R-3.6.2/bin/x64/* and *C:/Rtools/bin*
2. [Install Rtools](https://cran.r-project.org/bin/windows/Rtools/) for building R packages.
2. [download R-studio](https://rstudio.com/products/rstudio/download/)
3. Open R-studio and auto-detect R
4. Install additional packages

After installing R, change the path sequence so that packages install for all users:

```{r}
ls_spn_paths <- .libPaths()
print(ls_spn_paths)
ls_spn_paths <- c(ls_spn_paths[2], ls_spn_paths[1])
.libPaths(ls_spn_paths)
ls_spn_paths <- .libPaths()
print(ls_spn_paths)
```

# Install Packaging tools
install.packages(c("devtools", "pkgdown", "roxygen2", "bookdown", "knitr", "kableExtra", "formatR", "revealjs"))

# Install Statistics models
install.packages(c("AER", "minpack.lm"))
install.packages(c("quantreg"))

# for sublime r-ide
# install.packages("languageserver")

# Personal Packages
devtools::install_github("fanwangecon/REconTools")
devtools::install_github("fanwangecon/PrjOptiAlloc")

# Stata in Rmd
devtools::install_github("Hemken/Statamarkdown")
```

Temp Installs:
```
# 2020-10-19
# Temp install development version due to but
# https://github.com/rstudio/reticulate/issues/831
devtools::install_github("rstudio/reticulate")
```

## Update R on Computer

@@
